Manuscript 1 code

### 2024-09-11 manuscript_1_code.Rmd
Models and figures for figure 4 :call_me_hand:

# Functional Group Prep

@@

## SYNCHRONY CODE
```{r calculate syncrony}
# make er long
macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

@@

# CALCULATE SYNCHRONY CODE
lm_synchrony <- codyn::synchrony(df = macro_long_data,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "prop_cover",
                                   metric = "Loreau",
                                   replicate.var = "location")
```

>>>>

## prep

@@

# Figure 3

## model

```{r figure 3 model}
alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover)
summary(cover_mod)
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different

## figure 3 plot setup

```{r figure 3 setup}
diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm <- emmip(cover_mod,
                        habitat ~ richness,
                        at = list(richness = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects <- cover_emm %>%
  left_join(diversity_ranges_quad, by = c("habitat")) %>%
  filter(richness >= min_richness & richness <= max_richness)

alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

## figure 3 plot

```{r figure 3 plot}
(figure_3 <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro_with_cover,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

```{r figure 3 v2}
(figure_3_v2 <- figure_3 + facet_wrap(~habitat))
```
# Figure 4

## Synchrony setup
```{r calculate syncrony}
# make er long
macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

@@

# CALCULATE SYNCHRONY CODE
lm_synchrony <- codyn::synchrony(df = macro_long_data,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "prop_cover",
                                   metric = "Loreau",
                                   replicate.var = "location")

diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
  dplyr::group_by(location, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(macroalgal_prop_cover),
            cover_cv = CV(macroalgal_prop_cover),
            cover_stability = 1/CV(macroalgal_prop_cover))

diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs

dss_ranges<- diversity_stability_synchrony %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability),
            min_synchrony = min(synchrony_trans),
            max_synchrony = max(synchrony_trans))
```

## figure 4a model - richness stability

```{r richness synchrony mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
```

```{r figure 4a setup}

rich_stab_emm <- emmip(rich_stab_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects <- rich_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

## plot figure 4a

```{r figure 4a}
(figure_4a <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
```

## figure 4c model - richness synchrony

```{r richness synchrony mod}
rich_synch_mod <- glmmTMB(synchrony_trans ~ richness_mean*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
summary(rich_synch_mod)
hist(residuals(rich_synch_mod)) # looks good
plot(residuals(rich_synch_mod) ~ fitted(rich_synch_mod)) # same issue as always
r.squaredGLMM(rich_synch_mod) # 0.701, 0.775
emtrends(rich_synch_mod, pairwise ~ habitat, var = "richness_mean") # forereefs aren't different
```

```{r figure 4c setup}

rich_synch_emm <- emmip(rich_synch_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_synch_effects <- rich_synch_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

## plot figure 4c

```{r figure 4c}
(figure_4c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = synchrony_trans, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_synch_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_synch_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Synchrony\n", x = "Macroalgal Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
```
